---
title: 'Lab 3: Data Wrangling on Soccer Tournament Data'
date: 'July 9, 2020'
output:
  html_document: default
  pdf_document: default
---
```{r setup, message = F, include=FALSE}
library(tidyverse)
```

Soccer tournament data wrangling
--------------------------
Read the dataset of football games.
```{r}

 d<-read_csv("C:/Users/18747/Documents/github/R-course/data/results.csv",)
 head(d)
 
```

1. Select variables `date`, `home_team` and `away_team`.
```{r}
d%>%select(date)
d%>%select(home_team)
d%>%select(away_team)
```
2. Subset games with **Brazil** as the home team.
```{r}
d%>%filter(home_team=="Brazil")
```

3. Choose the games that Brazil won as the home team, and select variables `date`, `away_team` and `tournament`.
```{r}
d%>%filter(home_team=="Brazil" & home_score>away_score )%>%select(date,away_team,tournament)
```

4. Add the difference of goals, and an indicator variable called `goleada` for when the difference of goals is large,  and select what we did only for Brazil. **Hint: use `ifelse`**.
```{r}
c<-d%>%mutate(goleada=abs(home_score-away_score))%>%filter(home_team=="Brazil"|away_team=="Brazil")%>%arrange(goleada)%>%filter(goleada>5)
```

5. What was the largest difference in goals within these games?
```{r}
b<-c%>%arrange(desc(goleada))
b[1,]
```

6. The top 5 goleadas?
```{r}
b[c(1:5),]
```
7. Summary on goals scored by home teams, such as `mean` of `home_score` and `away_score`, `std`, using `group_by` and `summarise`
```{r}
d%>%group_by(home_team)%>%summarize(mean1=mean(away_score))
d%>%group_by(home_team)%>%summarize(mean2=mean(home_score))
```

8. Proportion of victories of **Brazil** on different tournaments against each opponent, for instance, **Argentina**.
```{r}
e<-d%>%filter((home_team=="Brazil"&away_team=="Argentina")|(away_team=="Brazil"&home_team=="Argentina"))%>%filter(home_team=="Argentina")

f<-e%>%filter(home_score<away_score)

g<-d%>%filter((home_team=="Brazil"&away_team=="Argentina")|(away_team=="Brazil"&home_team=="Argentina"))%>%filter(home_team=="Brazil")

h<-g%>%filter(home_score>away_score)

(dim(f)[1]+dim(h)[1])/(dim(g)[1]+dim(e)[1])
```
